---
title: "Creating a List"
---

## How to generate a list

We construct lists using the `list()` function. Each element of the list can be of a different type, and R stores them without coercion. We can even put vectors as element of a list. This is one of the defining characteristics of lists: they preserve the type and structure of their contents. 

```{r}
(l1 <- list(1, 2, 3))
(l2 <- list("bob", TRUE, 33, 42+3i))
```

We can even put vectors as element of a list.

```{r}
(l3 <- list(c(1,2),c(3,4)))
```

## Naming List Elements

While list elements are unnamed by default, it is often helpful to assign names to the elements, particularly when the list is intended to represent a structured record or object. Names can be provided directly when the list is created.

```{r}
(l4 <- list(name = "bob",
            alive = TRUE,
            age = 33,
            relationshipStatus = 42+3i))

```

This named list behaves much like a simple record or dictionary, with named fields that can be accessed directly. Alternatively, names can be assigned or modified after list creation using the `names()` function.

```{r}
names(l1) <- c("first", "second", "third")
l1
```






