---
title: "Conclusion"
format:
  html: 
    toc: false
    link-external-icon: false
    link-external-newwindow: false
---

```{r include = FALSE}
source(here::here("R", "buttons.R"))
```

By completing these four tutorials, you've built a solid foundation in R's core data types —vectors, matrices, lists, data frames, and factors. These structures are more than just tools; they are the building blocks of everything you'll do in R, from simple data manipulation to complex modeling and function development.

You should now have the confidence to recognize and work with each structure effectively, make informed choices about how to represent your data, and apply consistent techniques for accessing and transforming information. These core skills are essential not only for writing clear and reliable R code, but also for understanding how R uses the data in different ways.


## Next Steps 

With this knowledge in place, you're ready to shift focus from what R works with to how you work in R. In the next module, we’ll explore workflows—covering scripts, working directories, R projects, paths, and coding style—to help you build reproducible, organized, and efficient R projects from the ground up.

- Here, we're linking to the *Workflow* module. (Wrong link)

```{r}
#| echo: false
#| results: asis
next_module_button("https://www.openstatprog.com/first-steps/")
```
