---
title: "Conclusion"
format:
  html: 
    toc: false
    link-external-icon: true
    link-external-newwindow: true
    link-external-filter: ^(?:http:|https:)\/\/www\.openstatprog\.com
---

```{r include = FALSE}
source(here::here("R", "buttons.R"))
```

By completing these tutorials, you've built a solid foundation in understanding five of the most ubiquitous and useful
data types in R:

- Vectors
- Matrices
- Lists
- Data frames
- Factors

## Additional Reading

For more information on the topics we've covered in this module, the following chapters from Hadley Wickham's *R for
Data Science* are excellent resources.

- [Vectors](https://r4ds.had.co.nz/vectors.html) in the [First Edition](https://r4ds.had.co.nz/index.html).
- [Factors](https://r4ds.hadley.nz/factors.html) in the [Second Edition](https://r4ds.hadley.nz/index.html).

### Advanced Reading

For a much more technical deep-dive into the topics we've covered in this module, check out the following two chapters
from Hadley Wickham's [Advanced R](https://adv-r.hadley.nz/index.html).

- [Vectors](https://adv-r.hadley.nz/vectors-chap.html)
- [Subsetting](https://adv-r.hadley.nz/subsetting.html)

## Next Steps 

Now that you have a good introduction to the fundamental elements of the R statistical programming language, you're
ready to start thinking about how to best organize your data analysis projects. In the next module, we’ll explore
different aspects of effective R workflows--—scripts, working directories, R projects, file paths, and coding style—--to
help you create reproducible, organized, and efficient data analysis projects.

```{r}
#| echo: false
#| results: asis

# next_module_button("https://www.openstatprog.com/workflow/index.html")
next_module_button("in_progress.html")
```
