---
title: "Manipulating Matrices"
---

## Accessing Matrix Elements

As with vectors, we can access and assign values within a matrix using the square brackets operator, `[]`, but with one
key difference: matrices require two indices, one to define the row selection and one to define the column selection.

```{r}
(m1 <- matrix(1:12, 3, 4))

# Select the first two rows and the first three columns, then print the selected
# elements as a matrix
m1[1:2, 1:3]

# Select the first and third rows and second and third column, then save the
# selection as a new matrix called `m2`
(m2 <- m1[c(1, 3), 2:3])
```

## Modifying Matrix Elements

Editing the value of matrix elements works the same way it did for vectors. We , use the `[]` operator to select the
values we want to overwrite, and assign new values for the selected elements using the ordinary assignment operator,
`<-`.

```{r}
(m3 <- matrix(1, 3, 3))

# Replace the element in the first row and second column with 33
m3[1, 2] <- 33
m3

# Replace all elements in the range covered by the second and third rows and the
# first and third columns with 44 
m3[2:3, c(1, 3)] <- 44
m3
```
