---
title: "Creating Matrices"
---

The most direct way to create a new matrix is the `matrix()` function.

```{r}
(m1 <- matrix(data = 1, nrow = 3, ncol = 3))

```

If we inspect the object, we’ll see it now has an *attribute*, `dim`. The attribute `dim` is a two element vector, in
which the first element shows the number of rows and the second element the number of columns.

```{r}
# Create a numeric vector for comparison
y1 <- c(1, 2, 3)

# Basic vectors don't have attributes
attributes(y1)

# Matrices must have at least a 'dim' attribute
attributes(m1)
```

## The Matrix/Vector Relation

I'm not being facetious when I say that a matrix is just a vector with a `dim` attribute. In fact, we can convert a
vector to a matrix simply by adding a `dim` attribute to the vector.

```{r}
# y1 is an ordinary numeric vector
class(y1)
is.vector(y1)
is.matrix(y1)

# m1 is a matrix (which is a special case of an 'array')
class(m1)
is.vector(m1)
is.matrix(m1)

# Add a 'dim' attribute to y1
attr(y1, "dim") <- c(3, 1)

# Now, y1 looks like a 3x1 matrix
y1

# Indeed, y1 is now a matrix, as far as R is concerned
class(y1)
attributes(y1)
is.vector(y1)
is.matrix(y1)
```

Similarly, we can convert a matrix to a vector by removing the `dim` attribute from the matrix.

```{r}
# Setting the 'dim' attribute of m1 to NULL effectively removes that attribute
attr(m1, "dim") <- NULL

# Now, m1 looks like a vector
m1

# And R agrees: m1 is now a vector, not a matrix
class(m1)
is.vector(m1)
is.matrix(m1)
```

## Element Ordering

By default, R fills matrices column-wise (i.e., using *column-major order*): the first column is filled top to bottom,
then the second column, and so on.

```{r}
matrix(1:9, 3, 3)
```

If we want to fill the matrix row-by-row instead (i.e., using *row-major order*), we can use the `byrow = TRUE` argument.

```{r}
matrix(1:9, 3, 3, byrow = TRUE)
```
