---
title: "Matrix Operations"
---

```{r, echo = FALSE, include = FALSE}
set.seed(235711)
```

## Size

The `length()` function works with matrices, but the results may not be that interesting. When applied to matrices,
`length()` returns the number of cells in the matrix.

```{r}
(m1 <- matrix(1:9, 3, 3))
length(m1)
```

To get the dimensions of the matrix, we use `dim()`.

```{r}
dim(m1)
```

## Arithmetic

As with vectors, default arithmetic with R matrices works *element-wise*. R performs the requested operation on each
pair of corresponding entries in the two matrices. 

```{r}
(m2 <- matrix(sample(-5:5, 9, TRUE), 3, 3))

m1 + m2
m2 - m1
m1 / m2
m1 * m2
```

## Matrix Algebra

If you're familiar with matrix algebra and/or have some affinity for programming languages that overload their
operators, the matrix arithmetic described above may seem very strange. Rather than overloading the standard operators,
R defines special functions for matrix algebraic operations. For example:

- `%*%`: Multiplication
- `t()`: Transposition
- `solve()`: Inversion

```{r}
# Multiply m1 and m2 using true matrix multiplication
m1 %*% m2

# Transpose m1
t(m1)

# Invert m2
solve(m2)
```

## Recycling

Matrices usually obey R’s recycling rules. If you attempt to perform arithmetic between a matrix and a vector that has
fewer elements than the matrix, R will try to make the lengths match by recycling the elements from the vector.

```{r, warning = TRUE}
v1 <- 1:3
v2 <- 1:2

# Recycling the elements of 'v1' without a warning
m1 + v1
m1 * v1

# Recycling the elements of 'v2' with a warning
m1 + v2
m1 * v2

# Use recycling to fill a new matrix
matrix(42, 2, 2)
matrix(1:4, 4, 4)
matrix(1:3, 4, 4)

# Use recycling to match replacement values to all selected elements
m2[1:2, 3] <- -99
m2

m2[3, 1:2] <- c(88, 99)
m2
```

Sometimes, R won't apply recycling to accommodate our sloppiness, though. For example, we can't do any arithmetic with
matrices that have different dimensions.

```{r, error = TRUE}
(m3 <- matrix(c(1, 3), 2, 1))
m1 + m3
m1 * m3
```

We also can't do arithmetic between a matrix and vector that contains *more* elements than the matrix.

```{r, error = TRUE}
v3 <- 1:18
m1 + v3
m1 * v3
```

When it comes to overwriting matrix elements, R is especially picky about what size of vector it will use to overwrite a
selection of matrix elements.

- OK
   - Replacement length = Selection length
   - Replacement length cleanly divides selection length
- Not OK
   - Replacement length > Selection length
      - Still true when replacement length is a multiple of selection length
   - Replacement length *does not* cleanly divide the selection length

```{r, error = TRUE}
# Works: Replace 4 elements with length-4 vector
m1[1:2, 2:3] <- 1:4
m1

# Works: Replace 4 elements with length-2 vector
m1[1:2, 2:3] <- 1:2
m1

# Works: Replace 4 elements with length-1 vector
m1[1:2, 2:3] <- 42
m1

# Fails: Replace 4 elements with length-3 vector
m1[1:2, 2:3] <- 1:3

# Fails: Replace 4 elements with length-8 vector
m1[1:2, 2:3] <- 1:3
```

::: practice

## Practice

1. Create a 5x3 numeric matrix called 'myMat' wherein each column is equal to the vector 1:5.
1. Multiply each entry in 'myMat' by pi (i.e., the mathematical constant).

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}

```

## {{< var tab_title.hint >}}

HINT: The built-in R object `pi` contains the value of pi.

## {{< var tab_title.solution >}}

```{r}
(myMat <- matrix(1:5, 5, 3))

myMat * pi
```

::::

:::
