---
title: "Creating matrices"
---

## Working with Matrices in R

At their core, matrices in R are just atomic vectors with a special attribute: dimensions. When you add a `dim` attribute to a vector, R interprets it as a matrix. This means all the rules of vectors still apply—matrices simply add a structured, two-dimensional layout on top.

## Creating Matrices

The most common way to create a matrix is with the `matrix()` function. We can specify the number of rows and columns, and optionally the data to fill the matrix:

```{r}
(m1 <- matrix(data = 1, nrow = 3, ncol = 3))
```

If we inspect the object, we’ll see it now has additional attributes, `dim`:

```{r}
y1 <- c(1,2,3)

attributes(y1)
attributes(m1)
```

By default, R fills matrices column-wise. That means the first column is filled top to bottom, then the second column, and so on:

```{r}
(m2 <- matrix(1:9, 3, 3))
```

If we want to fill the matrix row by row instead, we can use the `byrow = TRUE` argument:

```{r}
(m3 <- matrix(1:9, 3, 3, byrow = TRUE))
```









