```{r, echo = FALSE, include = FALSE}

df <- data.frame(
  name = c("bob", "mike", "mark", "steve"),
  age = c(25, 23, 24, 26),
  height = c(180, 178, 169, 183),
  eyes = c("green", "green", "blue", "brown")
)

```

Question
========

Consider the following data frame, `df`.

```{r, echo = FALSE}
df
```

How can we extract the `height` column and return the selection as a *data frame*?

Answerlist
-----------
* `df[3]`
* `df["height"]`
* `df$height`
* `df[["height"]]`
* `df[[3]]`
* `df[ , 3]`
* `df[ , "height"]`
* `df[3, ]`
* `df["height", ]`
* `df[[2]]`
* `df[2:3]`

Solution
========

Answerlist
-----------
* Correct
* Correct
* Wrong: This command will return the `height` column as a numeric vector.
* Wrong: This command will return the `height` column as a numeric vector.
* Wrong: This command will return the `height` column as a numeric vector.
* Wrong: This command will return the `height` column as a numeric vector.
* Wrong: This command will return the `height` column as a numeric vector.
* Wrong: This command will return the third row of `df` as a $1 \times 4$ data frame.
* Wrong: This command will try to return a row called "height" and fail because there is no such row in `df`.
* Wrong: This command will return the `age` column as a numeric vector.
* Wrong: This command will return the `age` and `height` columns as a $4 \times 2$ data frame.

Meta-information
================
exname: return_df_column_as_df
extype: mchoice
exsolution: 11000000000
exshuffle: 5
