```{r, echo = FALSE, include = FALSE}

df <- data.frame(
  name = c("bob", "mike", "mark", "steve"),
  age = c(25, 23, 24, 26),
  height = c(180, 178, 169, 183),
  eyes = c("green", "green", "blue", "brown")
)

```

Question
========

Consider the following data frame, `df`.

```{r, echo = FALSE}
df
```

How can we extract the `name` column and return the selection as a *character vector*?

Answerlist
-----------
* `df$name`
* `df[["name"]]`
* `df[[1]]`
* `df[ , 1]`
* `df[ , "name"]`
* `df[1]`
* `df["name"]`
* `df[1, ]`
* `df["name", ]`
* `df[[2]]`
* `df[1:2]`

Solution
========

Answerlist
-----------
* Correct
* Correct
* Correct
* Correct
* Correct
* Wrong: This command will return the `name` column as a one-column data frame.
* Wrong: This command will return the `name` column as a one-column data frame.
* Wrong: This command will return the first row of `df` as a $1 \times 4$ data frame.
* Wrong: This command will try to return a row called "name" and fail because there is no such row in `df`.
* Wrong: This command will return the `age` column as a numeric vector.
* Wrong: This command will return the `name` and `age` columns as a $4 \times 2$ data frame.

Meta-information
================
exname: return_df_column_as_vector
extype: mchoice
exsolution: 11111000000
exshuffle: 5
