---
title: "Vector Elements"
---

## Accessing and Assigning Vector Elements

We can access or modify specific elements in a vector using square brackets `[]`. To retrieve the value at a specific position, we use the element’s index inside the brackets. Similarly, we can assign new values to individual elements by specifying their index.

```{r}
(y <- 1:5)
y[3]

y[1] <- 77
y
```


To modify multiple elements at once, we provide a vector of indices inside the brackets. 

```{r}
y[c(2, 4)] <- -22
y
```


::: {.callout-note}

Note that when subsetting, all indices must be placed inside a single pair of square brackets. Using a comma inside `[]` is reserved for multi-dimensional structures like matrices or data frames—and will result in an error when used with vectors.

:::

::: practice

## Practice

a. Create a vector with the numbers from 1 to 5
b. Substitute all the even numbers with 0

:::: {.panel-tabset}
## {{< var tab_title.question >}}

```{webr-r}

```

## {{< var tab_title.solution >}}

```{r}
(y <- 1:5)

y[c(2,4)] <- 0
y
```
::::

:::


