```{r, echo = FALSE, include = FALSE}
l1 <- list(name = "bob",
           age = 25,
           height = 180,
           eyes = "blue")
```

Question
========
Consider the following list, `l1`.

```{r, echo = FALSE}
l1
```

Which of the following expressions will change the value of the `eyes` slot in `l1` from "blue" to "green"?

Answerlist
----------
* `l1$eyes <- "green"`
* `l1[[4]] <- "green"`
* `l1[4] <- "green"`
* `l1$eyes <- "blue"`
* `l1$4 <- "green"`
* `l1$eyes <- green`
* `l1[4, ] <- "green"`
* `names(l1)[4] <- "green"`

Solution
========

Answerlist
----------
* Correct
* Correct
* Correct
* Wrong: This command doesn't actually change anything.
* Wrong: There is no slot named "4" in l1.
* Wrong: We're replacing the value of `l4$eyes` with the contents of the object `green`, not the character vector `"green"`.
* Wrong: l1 is a list, so we should not use matrix-style selection.
* Wrong: This command will change the name of the fourth slot, not the contents of that slot.

Meta-information
================
exname: change_list_slot
extype: mchoice
exsolution: 11100000
exshuffle: 5
