Question
========
```{r, echo = FALSE, include = FALSE}
l <- list(name = "bob",
          age = 25,
          height = 180,
          hair = "blue")
```

In the list `l` below there is an error, the information for the eyes color is wrongly saved as `hair`, which of this options will change the name of the element?

```{r, echo=FALSE}
l
```


Answerlist
----------
* `names(l)[4] <- "eyes"`
* `l$hair <- names("eyes")`
* `l[[4]] <- "eyes"`
* `l[4] <- "eyes"`
* `names(l)$hair <- "eyes"`

Solution
========
The correct way to change the name of the value is `names(l)[4] <- "eyes"`

Answerlist
----------
* Correct
* Wrong: `names("eyes")` is `NULL` since a character string doesn't have a name, so it will just erase the list element
* Wrong: This will change the value itself and not the name
* Wrong: This will change the value itself and not the name
* Wrong: This will return an error since the `names()` element is a vector and not a list

Meta-information
================
exname: Default Arguments Selection
extype: schoice
exsolution: 10000
exshuffle: 5
