---
title: "Creating factors"
---

## Creating and inspecting factors

We create factors in R using the `factor()` function. This function takes a vector of values (typically numeric or character) and converts it into a factor by assigning each unique value a level. The factor object stores the levels as labels and internally represents them as integers:

```{r}
(f1 <- factor(sample(1:3, 20, TRUE), labels = c("red", "yellow", "blue")))
```

Although a factor appears to display text labels, it is actually stored internally as an integer vector. We can verify this by inspecting its type and attributes:

```{r}
typeof(f1)       
attributes(f1)   
```

The levels define the possible categories, and every value in the factor must be one of these levels. The ordering of levels can also be important when working with ordered factors, although by default, factors are treated as unordered.
The labels associated with a factor are stored in its `levels` attribute:

```{r}
levels(f1)
```

::: practice 

## Practice

Create a length-20 factor with two levels = {"yes", "no"}.

:::: {.panel-tabset}
## {{< var tab_title.question >}}

```{webr-r}

```


## {{< var tab_title.solution >}}

```{r}
f_ex <- factor(sample(c(1,2), 20, TRUE), labels = c("yes", "no"))
```
::::

:::


