---
title: "Creating Lists"
---

## Generating Lists

We construct lists using the `list()` function. Lists place essentially no restrictions on the nature of their contents:

- A list slot can hold any valid R object (including other lists).
- The contents of different list slots do not need to have the same type (nor do they need to be different types).
- The contents of different list slots do not need to have the same---or otherwise conformable---sizes.

```{r}
(l1 <- list(1, 2, 3))
(l2 <- list("bob", TRUE, 33, 42+3i))
(l3 <- list(c(1, 2), c(3, 4)))
```


## Naming List Elements

We can name the elements of a list by defining the list slots via key-value pairs.

```{r}
l4 <- list(name = "bob",
           alive = TRUE,
           age = 33,
           relationshipStatus = 42+3i)
l4
```

We can use the `names()` function to access and modify the names of a list's slots.

```{r}
names(l4)
names(l1) <- c("first", "second", "third")
l1
```

::: practice

## Practice

Create a list to describe yourself. Include the following named elements in your list:

1. Your Name
1. Your Eye Color
1. Your Hair Color
1. Your Favorite Color

:::: {.panel-tabset}

## {{< var tab_title.question >}}

```{webr-r}

```

## {{< var tab_title.solution >}}

```{r}
myInfo <- list(name = "bob",
               eyes = "blue",
               hair = "black",
               color = "green")
myInfo
```

::::

:::
